---
title: "Hmk06"
author: "Ziyi Tao"
output: html_document
date: "2024-11-30"
---

Github : <https://github.com/MelDrawer/Stats506_Hmk6.git>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

### a. Estimate standard deviation

a.1 Average RF for each team:

```{r}
library(DBI)
library(tidyr)
library(dplyr)

# import data
lahman <- dbConnect(RSQLite::SQLite(), 
                    "D:/lahman_1871-2022.sqlite")
fielding <- dbGetQuery(lahman, "SELECT * FROM fielding")
```

```{r}
fielding %>%
filter(InnOuts !=0)%>%
group_by(teamID) %>%
mutate(RF = 3*(PO+A)/InnOuts)%>%
summarize(mean_RF = mean(RF, na.rm = TRUE))%>%
arrange(-mean_RF) -> mean_rf

head(mean_rf)
```

a.2 Estimate without parallel processing:

```{r}
# 1 drop InnOuts == 0
fielding %>%
  filter(InnOuts !=0)%>%
  mutate(RF = 3*(PO+A)/InnOuts) -> fielding_1

# 2 set boots numbers
boots <- 1000

# 3 stratified bootstrap function
#' boot
#' @param team_data
#' @return the mean RF of each bootstarpped sample of one team
boot <- function(team_data){
  sample_boot <- sample(team_data$RF, replace = TRUE) 
  boot_meanrf <- mean(sample_boot, na.rm = TRUE)
  return(boot_meanrf)
}


```

```{r}
# 4 without parallel processing
estimate_sd <- data.frame(team = unique(fielding_1$teamID), 
                           sd_boot = NA, meanrf_boot = NA)

for(id in unique(fielding_1$teamID)){
  for(i in 1:boots){
    boot_meanrf[i] = boot(fielding_1[fielding_1$teamID == id,])
  }
  estimate_sd[estimate_sd$team == id, "sd_boot"] =  sd(boot_meanrf)
  estimate_sd[estimate_sd$team == id, "meanrf_boot"] =  mean(boot_meanrf)
}

```

```{r}
head(estimate_sd)
```

a\. 3 Using parallel processing with parallel package

```{r}
library(parallelly)
library(parallel)
boots <- 1000


#' boot_n
#' @param team 
#' @param boots 
#' @return the standard deviation in bootstrap simulations and the estimated mean of rf given certain team
boot_n <- function(team, boots){
  boot_meanrf <- rep(NA, boots)
  team_data <- fielding_1[fielding_1$teamID == team,]
  for(i in 1:boots){
    sample_boot <- sample(team_data$RF, replace = TRUE) 
    boot_meanrf[i] <- mean(sample_boot, na.rm = TRUE)
  }
  return(c(sd(boot_meanrf), mean(boot_meanrf)))
}


cl <- makeCluster(8)
clusterExport(cl, c("fielding_1", "boot_meanrf", "boots"))
res_parallel <- parSapply(cl, 
                          unique(fielding_1$teamID), 
                          boot_n, boots)
stopCluster(cl)

```

```{r}
estimate_sd_parallel <- data.frame(sd_boot = res_parallel[1,], 
                    team = colnames(res_parallel), 
                    mearf_boot = res_parallel[2,])

head(estimate_sd_parallel)
```

a.4 Using future with future package

```{r}
library(future)

teams <- unique(fielding_1$teamID)

#' boot_teams
#' @param teams 
#' @param boots 
#' @return the estimated meanrf and standard deviation of meanrf for teams
boot_teams <- function(teams, boots){
  res <- as.data.frame(cbind(team = teams, sd_boot = NA, meanrf_boot = NA))
  for(i in teams){
    res[res$team == i, "sd_boot"] = boot_n(i, boots)[1]
    res[res$team == i, "meanrf_boot"] = boot_n(i, boots)[2]
  }
  return (res)
}

plan(multisession)
b1 %<-% boot_teams(teams[1:35], 1000)
b2 %<-% boot_teams(teams[36:70], 1000)
b3 %<-% boot_teams(teams[71:105], 1000)
b4 %<-% boot_teams(teams[106:140], 1000)
res_future <- rbind(b1, b2, b3 ,b4)
plan(sequential)


```

```{r}
head(res_future)
```

### b. Show final result

```{r}
# select top 10 teams
teams_top10 <- mean_rf[1:10,]

# without parallel
res_1 <- merge(teams_top10, estimate_sd, 
               by.y = "team", by.x = "teamID")
res_1[order(res_1$mean_RF, decreasing = TRUE),]
```

```{r}
# with parallel
res_2 <- merge(teams_top10, estimate_sd_parallel, 
               by.y = "team", by.x = "teamID")
res_2[order(res_2$mean_RF, decreasing = TRUE),]
```

```{r}
# future
res_3 <- merge(teams_top10, res_future, 
               by.y = "team", by.x = "teamID")
res_3[order(res_3$mean_RF, decreasing = TRUE),]
```

### c. Discussion

As is shown in the result, bootstrap without any parallel processing has the longest running time 945.57s while parallel processing using 8 clusters has the shortest running time 7.20s, this shows that parallel processing can greatly increase the running speed. When we use future package, the running time is also greatly decreased.

```{r}
# with out parallel
system.time({
  estimate_sd <- data.frame(team = unique(fielding_1$teamID), 
                           sd_boot = NA, meanrf_boot = NA)

  for(id in unique(fielding_1$teamID)){
    for(i in 1:boots){
      boot_meanrf[i] = boot(fielding_1[fielding_1$teamID == id,])
    }
    estimate_sd[estimate_sd$team == id, "sd_boot"] =  sd(boot_meanrf)
    estimate_sd[estimate_sd$team == id, "meanrf_boot"] =  mean(boot_meanrf)
  }
})


```

```{r}
# parallel
system.time({
  cl <- makeCluster(8)
  clusterExport(cl, c("fielding_1", "boot_meanrf", "boots"))
  res_parallel <- parSapply(cl, 
                            unique(fielding_1$teamID), 
                            boot_n, boots)
  stopCluster(cl)
  
})
```

```{r}
system.time({
  plan(multisession)
  b1 %<-% boot_teams(teams[1:35], 1000)
  b2 %<-% boot_teams(teams[36:70], 1000)
  b3 %<-% boot_teams(teams[71:105], 1000)
  b4 %<-% boot_teams(teams[106:140], 1000)
  res_future <- rbind(b1, b2, b3 ,b4)
  plan(sequential)
})
```
